---
title: "Red blood cell distribution width to platelet ratio: A novel noninvasive index aids in the diagnosis of sepsis in neonatal foals"
subtitle: Code notebook by Gabriela Novo de Oliveira and McCaide Wooten
output: 
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_depth: 4
    depth: 2
    number_sections: false
    toc_float:
      collapsed: true
---

# Dataset overview

### Loading packages

```{r}
library(plyr)
library(dplyr)
library(lubridate)
library(data.table)
library(forcats)
library(ggplot2)
library(knitr)
library(lubridate)
library(purrr)
library(rmarkdown)
library(stringr)
library(tibble)
library(tidyr)
library(MASS)
library(stats)
library(survival)
library(labelled)
library(DescTools)
```

### **Loading the dataset ('dataset_modified')**

```{r}
rdw_raw <- read.csv("~/dataset_modified.csv")
```

### **Dimension of the dataset**

```{r}
dim(rdw_raw)
```

### **Structure of the dataset**

```{r}
str(rdw_raw)
```

### **Variable names**

```{r}
names(rdw_raw)
```

### Summary of variables

```{r}
summary(rdw_raw)
```

### **Labeling variables**

```{r}
var_label(rdw_raw$discharge_status) <- 'Discharge status (alive, euth, died)'
var_label(rdw_raw$age) <- 'Age at bloodsampling (in hours)'
var_label(rdw_raw$group) <- 'Group according to sepsis score'
var_label(rdw_raw$premature) <- 'Premature (yes/no)'
var_label(rdw_raw$gestational_age) <- 'Gestational age (in days)'
var_label(rdw_raw$blood_culture_done) <- 'Blood culture: yes (1) or no(2)'
var_label(rdw_raw$blood_culture_result) <- 'Blood culture result: positive (1) or negative (2)'

```

### **Changing variable types**

```{r}
rdw_raw$R <- as.numeric(rdw_raw$R)
rdw_raw$P <- as.numeric(rdw_raw$P)
rdw_raw$R <- as.numeric(rdw_raw$R)
rdw_raw$creatinine <- as.numeric(rdw_raw$creatinine)
rdw_raw$CRT <- as.numeric(rdw_raw$CRT)
```

### **Looking at the first rows of the dataset**

```{r}
head(rdw_raw)
```

### Creating subsets

```{r}
rdw_healthy_group <- rdw_raw %>%
  filter(group %in% c('Healthy'))
dim(rdw_healthy_group)

rdw_sns_group <- rdw_raw %>%
  filter(group %in% c('SNS'))
dim(rdw_sns_group)

rdw_septic_group <- rdw_raw %>%
  filter(group %in% c('Septic'))
dim(rdw_septic_group)

rdw_subset_sick <- rdw_raw %>%
  filter(group %in% c('SNS','Septic'))

rdw_survivors <- rdw_raw %>%
  filter(survival_status %in% c('Survivor'))

rdw_non_survivors <- rdw_raw %>%
  filter(survival_status %in% c('Non-Survivor'))
```

### Kruskal-Wallis test to compare medians between groups

```{r}
kruskal.test(baso~group,data=rdw_raw) 
kruskal.test(creatinine~group,data=rdw_raw) 
kruskal.test(CRT~group,data=rdw_raw) 
kruskal.test(eos~group,data=rdw_raw) 
kruskal.test(glu~group,data=rdw_raw) 
kruskal.test(P~group,data=rdw_raw) 
kruskal.test(HCT~group,data=rdw_raw) 
kruskal.test(Hg~group,data=rdw_raw) 
kruskal.test(lactate~group,data=rdw_raw) 
kruskal.test(lymph~group,data=rdw_raw) 
kruskal.test(MCH~group,data=rdw_raw) 
kruskal.test(MCHC~group,data=rdw_raw) 
kruskal.test(MCV~group,data=rdw_raw) 
kruskal.test(mono~group,data=rdw_raw) 
kruskal.test(neut~group,data=rdw_raw) 
kruskal.test(platelets~group,data=rdw_raw) 
kruskal.test(RBC~group,data=rdw_raw) 
kruskal.test(RDW~group,data=rdw_raw) 
kruskal.test(R~group,data=rdw_raw) 
kruskal.test(RPR~group,data=rdw_raw) 
kruskal.test(sepsis_score~group,data=rdw_raw) 
kruskal.test(T~group,data=rdw_raw) 
kruskal.test(WBC~group,data=rdw_raw) 
kruskal.test(WBC~group,data=rdw_raw) 


```

### Medians for each group

#### Healthy group

```{r}
MedianCI(rdw_healthy_group$baso, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_healthy_group$creatinine, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_healthy_group$CRT, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_healthy_group$eos, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_healthy_group$glu, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_healthy_group$P, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_healthy_group$HCT, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_healthy_group$Hg, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_healthy_group$lactate, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_healthy_group$lymph, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_healthy_group$MCH, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_healthy_group$MCHC, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_healthy_group$MCV, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_healthy_group$mono, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_healthy_group$neut, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_healthy_group$platelets, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_healthy_group$RBC, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_healthy_group$RDW, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_healthy_group$R, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_healthy_group$RPR, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_healthy_group$sepsis_score, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_healthy_group$T, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_healthy_group$WBC, conf.level=0.95, na.rm=TRUE)

```

#### Sick non-septic (SNS) group

```{r}
MedianCI(rdw_sns_group$baso, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_sns_group$creatinine, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_sns_group$CRT, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_sns_group$eos, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_sns_group$glu, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_sns_group$P, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_sns_group$HCT, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_sns_group$Hg, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_sns_group$lactate, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_sns_group$lymph, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_sns_group$MCH, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_sns_group$MCHC, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_sns_group$MCV, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_sns_group$mono, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_sns_group$neut, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_sns_group$platelets, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_sns_group$RBC, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_sns_group$RDW, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_sns_group$R, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_sns_group$RPR, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_sns_group$sepsis_score, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_sns_group$T, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_sns_group$WBC, conf.level=0.95, na.rm=TRUE)
```

#### Septic group

```{r}
MedianCI(rdw_septic_group$baso, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_septic_group$creatinine, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_septic_group$CRT, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_septic_group$eos, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_septic_group$glu, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_septic_group$P, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_septic_group$HCT, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_septic_group$Hg, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_septic_group$lactate, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_sns_group$lymph, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_septic_group$MCH, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_septic_group$MCHC, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_septic_group$MCV, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_septic_group$mono, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_septic_group$neut, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_septic_group$platelets, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_septic_group$RBC, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_septic_group$RDW, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_septic_group$R, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_septic_group$RPR, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_septic_group$sepsis_score, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_septic_group$T, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_septic_group$WBC, conf.level=0.95, na.rm=TRUE)
```

## Testing the data for normality

### Q-Q plots

```{r}
qqPlot(rdw_raw$RDW,main='RDW')
qqPlot(rdw_raw$RPR,main='RPR')
```

### Histograms

```{r}
hist(rdw_raw$RDW)
hist(rdw_raw$RPR)
```

### Shapiro-Wilk test for normality

H0: data are normally distributed

Therefore, for p-values \< 0.05, we reject the null and conclude the data are NOT normally distributed

```{r}
shapiro.test(rdw_raw$RDW)
shapiro.test(rdw_raw$RPR)
```

All of the continuous variables yielded a p-value \<0.05, thus leading us to reject the null

### Testing data transformation

### Log transformation

```{r}
logRDW <- log(rdw_raw$RDW)
logRPR <- log(rdw_raw$RPR)

par(mfcol=c(2,2))
qqPlot(logRDW)
qqPlot(logRPR)
hist(logRDW)
hist(logRPR)

shapiro.test(logRDW) 
shapiro.test(logRPR) 
```

### Square-root transformation

```{r}
sqrtRDW <- sqrt(rdw_raw$RDW)
sqrtRPR <- sqrt(rdw_raw$RPR)

par(mfcol=c(2,2))
qqPlot(sqrtRDW)
qqPlot(sqrtRPR)
hist(sqrtRDW)
hist(sqrtRPR)

shapiro.test(sqrtRDW) 
shapiro.test(sqrtRPR) 
```

Log transformation seems like the best option, therefore it will be used for further analysis

### Log transformed dataset

```{r}
rdw_log <- rdw_raw %>%
  mutate(logRDW = log(rdw_raw$RDW)) %>%
  mutate(logRPR = log(rdw_raw$RPR)) 
```

Log transforming the data improved normality, but the data still doesn't follow a normal distribution. Therefore, non-parametric tests will be conducted.

## Kruskal-Wallis test

Kruskal-Wallis test is used to determine whether or not there is a statistically significant difference between the medians of three or more independent groups.

H0: the mean ranks of the groups are the same

**Kruskal-Wallis test for RDW:**

```{r}
kruskal.test(RDW~group,data=rdw_log) 
  #chi2 3.5677
  #p-value 0.168 (not significant)
kruskal.test(RDW~survival_status,data=rdw_log) 
  #chi2 1.2161
  #p-value 0.2701 (not significant)
wilcox.test(RDW~sex,data=rdw_log) 
  # W=12208
  #p-value 0.6974 (not significant)


kruskal.test(RDW~group,data=rdw_subset_sick) 
  #chi2 0.41788
  #p-value 0.518 (not significant)
wilcox.test(RDW~sex,data=rdw_subset_sick) 
  #W 6235
  #p-value 0.3349 (not significant)
kruskal.test(RDW~survival_status,data=rdw_log) 
  #chi2 1.2161
  #p-value 0.2701 (not significant)


wilcox.test(RDW~sex,data=rdw_septic_group) 
  # W=728.5
  #p-value 0.3349 (not significant)
kruskal.test(RDW~survival_status,data=rdw_septic_group) 
  #chi2 0.75883
  #p-value 0.3837 (not significant)

```

**Kruskal-Wallis test for RPR per categories**

```{r}
kruskal.test(RPR~group,data=rdw_log) 
  #chi2 14.703
  #p-value 0.0006417 (statistically significant)
wilcox.test(RPR~sex,data=rdw_log) 
  #W = 9920.5
  #p-value 0.001393 (statitiscally significant)
kruskal.test(RPR~survival_status,data=rdw_log) 
  #chi2 4.9995
  #p-value 0.02536 (statistically significant)


kruskal.test(RPR~group,data=rdw_subset_sick) 
  #chi2 8.4982
  #p-value 0.003555 (statistically significant)
wilcox.test(RPR~sex,data=rdw_subset_sick) 
  # W=4993
  #p-value 0.0009401 (statitiscally significant)
kruskal.test(RPR~survival_status,data=rdw_subset_sick) 
  #chi2 3.035
  #p-value 0.08149 (not statistically significant)


wilcox.test(RPR~sex,data=rdw_septic_group) 
  # W=683.5
  #p-value 0.1677 (not significant)
kruskal.test(RPR~survival_status,data=rdw_septic_group) 
  #chi2 0.0017533
  #p-value 0.9666 (not significant)
```

### Correlation tests for RDW\~continuous variables

```{r}

#correlation between RDW and days of hospitalization
cor.test(rdw_log$RDW,rdw_log$days_of_hospitalization,method=c("spearman"))
  #p-value 0.6813 (not significant)
  #rho 0.02315623 

#correlation between RDW and age
cor.test(rdw_log$RDW,rdw_log$age,method=c("spearman"))
  #p-value 0.6275 (not significant)
  #rho 0.02744654

#correlation between RDW and sepsis score
cor.test(rdw_log$RDW,rdw_log$sepsis_score,method=c("spearman"))
  #p-value 0.3167 (not significant)
  #rho 0.05641121




#correlation between RDW and days of hospitalization among the sick animals
cor.test(rdw_subset_sick$RDW,rdw_subset_sick$days_of_hospitalization,method=c("spearman"))
  #p-value 0.6578 (not significant)
  #rho -0.02625683 

#correlation between RDW and age among the sick animals
cor.test(rdw_subset_sick$RDW,rdw_subset_sick$age,method=c("spearman"))
  #p-value 0.6353 (not significant)
  #rho 0.03143857 

#correlation between RDW and sepsis score among the sick animals
cor.test(rdw_subset_sick$RDW,rdw_subset_sick$sepsis_score,method=c("spearman"))
  #p-value 0.6339 (not significant)
  #rho -0.0314312


#correlation between RDW and days of hospitalization among the septic animals
cor.test(rdw_septic_group$RDW,rdw_septic_group$days_of_hospitalization,method=c("spearman"))
  #p-value 0.8758 (not significant)
  #rho 0.01741893 

#correlation between RDW and age among the septic animals
cor.test(rdw_septic_group$RDW,rdw_septic_group$age,method=c("spearman"))
  #p-value 0.6736 (not significant)
  #rho 0.04721007 

#correlation between RDW and sepsis score among the septic animals
cor.test(rdw_septic_group$RDW,rdw_septic_group$sepsis_score,method=c("spearman"))
  #p-value 0.03211 (statistically significant)
  #rho 0.2354876


```

### Correlation tests for RPR\~continuous variables

```{r}
#correlation between RPR and days of hospitalization
cor.test(rdw_log$RPR,rdw_log$days_of_hospitalization,method=c("spearman"))
  #p-value 0.009524 (statistically significant)
  #rho 0.1454172

#correlation between RPR and age
cor.test(rdw_log$RPR,rdw_log$age,method=c("spearman"))
  #p-value 0.0001636 (statistically significant)
  #rho 0.210829

#correlation between RPR and sepsis score
cor.test(rdw_log$RPR,rdw_log$sepsis_score,method=c("spearman"))
  #p-value 0.0001538 (statistically significant)
  #rho 0.211014



#correlation between RPR and days of hospitalization among the sick animals
cor.test(rdw_subset_sick$RPR,rdw_subset_sick$days_of_hospitalization,method=c("spearman"))
  #p-value 0.03146 (statistically significant)
  #rho 0.1412877

#correlation between RPR and age among the sick animals
cor.test(rdw_subset_sick$RPR,rdw_subset_sick$age,method=c("spearman"))
  #p-value 0.0002652 (statistically significant)
  #rho 0.238309

#correlation between RPR and sepsis score among the sick animals
cor.test(rdw_subset_sick$RPR,rdw_subset_sick$sepsis_score,method=c("spearman"))
  #p-value 0.001241 (statistically significant)
  #rho 0.2107709


#correlation between RPR and days of hospitalization among the septic animals
cor.test(rdw_septic_group$RPR,rdw_septic_group$days_of_hospitalization,method=c("spearman"))
  #p-value 0.3538 (not significant)
  #rho 0.1030722

#correlation between RPR and age among the septic animals
cor.test(rdw_septic_group$RPR,rdw_septic_group$age,method=c("spearman"))
  #p-value 0.0008877 (statistically significant)
  #rho 0.3601889

#correlation between RPR and sepsis score among the septic animals
cor.test(rdw_septic_group$RPR,rdw_septic_group$sepsis_score,method=c("spearman"))
  #p-value 0.7791 (not significant)
  #rho 0.03125007

```

### Correlation tests for sepsis_score\~continuous variables

```{r}
#correlation between sepsis_score and days of hospitalization
cor.test(rdw_log$sepsis_score,rdw_log$days_of_hospitalization,method=c("spearman"))
  #p-value 1.628e-09 (statistically significant)
  #rho 0.330485

#correlation between sepsis_score and age
cor.test(rdw_log$sepsis_score,rdw_log$age,method=c("spearman"))
  #p-value 0.5468 (not significant)
  #rho 0.03408019


```

## Box-plots for significant differences between groups

```{r}
##### BOX PLOTS FOR RPR #####

### RPR AMONG ALL ###
#box plot RPR by group
rdw_log$group <- factor(rdw_log$group, levels=c('Healthy','SNS','Septic'))

boxplot(rdw_log$RPR~rdw_log$group, outline=FALSE, xlab='Group', ylab='RPR', main='RPR by group among all', col='darkgreen')
  

#box plot RPR by sex
boxplot(rdw_log$RPR~rdw_log$sex, outline=FALSE, xlab='Sex', ylab='RPR',main='RPR by sex among all', col='yellow',names=c('Female','Male'))


# box plot RPR by survival status
rdw_log$survival_status <- factor(rdw_log$survival_status, levels=c('Survivor','Non-Survivor'))

boxplot(rdw_log$RPR~rdw_log$survival_status, outline=FALSE, xlab='Discharge status', ylab='RPR', main='RPR by survival status among all', col='steelblue')

### RPR AMONG SICK ###
#box plot RPR by group among sick
boxplot(rdw_subset_sick$RPR~rdw_subset_sick$group, outline=FALSE, xlab='Group', ylab='RPR', main='RPR by group among sick', col='darkgreen')

#box plot RPR by sex among sick
boxplot(rdw_subset_sick$RPR~rdw_subset_sick$sex, outline=FALSE, xlab='Sex', ylab='RPR', main='RPR by sex among sick', col='yellow')
  


```

```{r}
par(mfcol=c(1,3))
#box plot RPR by group
boxplot(rdw_log$RPR~rdw_log$group, outline=FALSE, xlab='Group', ylab='RPR', main='RPR by group', names = c('Healthy (n = 85)', 'SNS (n = 149)', ' Septic (n = 83)'))

#box plot RPR by sex
boxplot(rdw_log$RPR~rdw_log$sex, outline=FALSE, xlab='Sex', ylab='RPR', main='RPR by sex',names=c('Female (n = 150)','Male (n = 167)'))

# box plot RPR by survival status
boxplot(rdw_log$RPR~rdw_log$survival_status, outline=FALSE, xlab='Discharge status', ylab='RPR', main='RPR by survival status',names=c('Survivor (n = 278)','Non-Survivor (n = 39)'))
```

## Scatter-plots

```{r}
### PLOT AMONG ALL ###
par(mfcol=c(4,2))

#scatter plot RPR x age with color for survival_status and shape for group
ggplot(rdw_log, aes(x = age, y = RPR)) +
  geom_point(aes(color=survival_status,shape=group),position = position_jitter(width=0.3))+
  stat_smooth(method='glm',se=TRUE)+
  ylim(0,0.4)


#scatter plot RPR x days_of_hospitalization with color for survival_status and shape for group
ggplot(rdw_log, aes(x = days_of_hospitalization, y = RPR)) +
  geom_point(aes(color=survival_status,shape=group),position = position_jitter(width=0.3))+
  stat_smooth(method='glm',se=TRUE)+
  xlim(0,44)+
  ylim(0,0.4)

#scatter plot RPR x sepsis score with color for survival_status and shape for group
ggplot(rdw_log, aes(x = sepsis_score, y = RPR)) +
  geom_point(aes(color=survival_status,shape=group),position = position_jitter(width=0.3))+
  stat_smooth(method='glm',se=TRUE)+
  ylim(0,0.4)


### PLOTS AMONG SICK ###

#scatter plot RPR x age with color for survival_status and shape for group
ggplot(rdw_subset_sick, aes(x = age, y = RPR)) +
  geom_point(aes(color=survival_status,shape=group),position = position_jitter(width=0.3))+
  stat_smooth(method='glm',se=TRUE)+
  ylim(0,0.4)

#scatter plot RPR x days_of_hospitalization with color for survival_status and shape for group
ggplot(rdw_subset_sick, aes(x = days_of_hospitalization, y = RPR)) +
  geom_point(aes(color=survival_status,shape=group),position = position_jitter(width=0.3))+
  stat_smooth(method='glm',se=TRUE)+
  xlim(0,44)+
  ylim(0,0.4)

#scatter plot RPR x sepsis score with color for survival_status and shape for group
ggplot(rdw_subset_sick, aes(x = sepsis_score, y = RPR)) +
  geom_point(aes(color=survival_status,shape=group),position = position_jitter(width=0.3))+
  stat_smooth(method='glm',se=TRUE)+
  ylim(0,0.4)



### PLOTS AMONG SEPTIC ###

#scatter plot RPR x age with color for survival_status 
ggplot(rdw_septic_group, aes(x = sepsis_score, y = RPR)) +
  geom_point(aes(color=survival_status),position = position_jitter(width=0.3))+
  stat_smooth(method='glm',se=TRUE)+
  ylim(0,0.4)

#scatter plot RDW x sepsis_score with color for survival_status
ggplot(rdw_septic_group, aes(x = sepsis_score, y = RDW)) +
  geom_point(aes(color=survival_status),position = position_jitter(width=0.3))+
  stat_smooth(method='glm',se=TRUE)
```

# Model building

## Missing data

Checking for missing data for each possible covariate

```{r}
for(i in names(rdw_log)){
print (c(i, sum(is.na((rdw_log[i])))))}
```

## Correlation between variables

Plotting the correlation between variables (numeric and categorical)

```{r}
head(rdw_log)
```

```{r}
correlation_plot  <- plot(rdw_log[c(6,12,28:34,36,46,50,52:56,59:61,64)])

correlation_plot
```

Finding which quantitative variables (covariates) are highly correlated with each:

```{r}
covariates = names(rdw_log)[(names(rdw_log) %in% c('days_of_hospitalization','age','RBC','Hg','HCT','MCV','MCH','MCHC','RDW','WBC','neut','lymph','mono','eos','baso','platelets','TPP','fib','RPR','creatinine'))]

covariates
```

```{r}
cor(rdw_log[,covariates])
```

Correlation matrix:

```{r}
cormat <- round(cor(drop_na(rdw_log[,covariates])),2)

cormat
```

Plotting the correlation matrix:

```{r}
melted_cormat <- melt(cormat)
head(melted_cormat)
```

```{r}
ggplot(data=melted_cormat, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile()
```

Making the plot above more readable:

```{r}
# reordering the correlation matrix
reorder_cormat <- function(cormat){
  # Use correlation between variables as distance
  dd <- as.dist((1-cormat)/2)
  hc <- hclust(dd)
  cormat <-cormat[hc$order, hc$order]
}

cormat <- reorder_cormat(cormat)

# Get upper triangle of the correlation matrix
get_upper_tri <- function(cormat){
  cormat[lower.tri(cormat)]<- NA
  return(cormat)
}

## remove the upper half of the matrix
upper_tri <- get_upper_tri(cormat)

melted_cormat <- melt(upper_tri, na.rm = TRUE)

# Creating a ggheatmap
ggheatmap <- ggplot(melted_cormat, aes(Var2, Var1, fill = value))+
  geom_tile(color = "white")+
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1,1), space = "Lab", 
                       name="") +
  theme_minimal()+ # minimal theme
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 12, hjust = 1))+
  coord_fixed()

# Printing the heatmap
ggheatmap + 
  geom_text(aes(Var2, Var1, label = value), color = "black", size = 4) +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.major = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.ticks = element_blank(),
    legend.justification = c(1, 0),
    legend.position = c(0.6, 0.7),
    legend.direction = "horizontal")+
  guides(fill = guide_colorbar(barwidth = 7, barheight = 1,
                               title.position = "top", title.hjust = 0.5))
```

We will keep a close eye on covariates that have correlation coefficient higher than 70%

We decided to drop Hg, RBC, **MCV, MCH** (highly correlated) - let's return to this after univariate analysis

## Single main effect plots

### Among all animals (healthy, SNS and septic)

```{r}
variables = names(rdw_log)[(names(rdw_log) %in% c('days_of_hospitalization','discharge_status','patient_breed','sex','age','group','HCT','MCHC','RDW','anisocytosis','WBC','neut','lymph','mono','eos','baso','platelets','TPP','fib','RPR','creatinine'))]

rdw_log$surv = factor(rdw_log$survival_status)

getBetaAndSe <- function(covar){
  coefs <- glm(surv~scale(rdw_log[[covar]]),data=rdw_log,family='binomial') %>% 
    summary() %>% coef()
  data.frame(variable = covar, 
             estimate = coefs[2,1], 
             se = coefs[2,2], 
             p = coefs[2,4])
}

beta.table <- adply(t(covariates), 2, getBetaAndSe) %>% 
  arrange(estimate) %>% 
  mutate(low = estimate - 2*se, 
         high = estimate + 2*se,
         variable = reorder(variable, order(estimate)))

ggplot(beta.table, aes(y=estimate, x=variable, ymin=low, ymax=high, color = (p < 0.05))) + 
  geom_pointrange() +  
  geom_hline(yintercept = 0, col="darkgrey", lty = 3, lwd=2) + 
  coord_flip() + theme_few() + ggtitle("Single main effect plot")

head(rdw_log)
```

### Among septic animals

```{r}
variables = names(rdw_septic_group)[(names(rdw_septic_group) %in% c('days_of_hospitalization','discharge_status','patient_breed','sex','age','group','HCT','MCHC','RDW','anisocytosis','WBC','neut','lymph','mono','eos','baso','platelets','TPP','fib','RPR','creatinine'))]

rdw_septic_group$surv = factor(rdw_septic_group$survival_status)

getBetaAndSe <- function(covar){
  coefs <- glm(surv~scale(rdw_septic_group[[covar]]),data=rdw_septic_group,family='binomial') %>% 
    summary() %>% coef()
  data.frame(variable = covar, 
             estimate = coefs[2,1], 
             se = coefs[2,2], 
             p = coefs[2,4])
}

beta.table <- adply(t(covariates), 2, getBetaAndSe) %>% 
  arrange(estimate) %>% 
  mutate(low = estimate - 2*se, 
         high = estimate + 2*se,
         variable = reorder(variable, order(estimate)))

ggplot(beta.table, aes(y=estimate, x=variable, ymin=low, ymax=high, color = (p < 0.05))) + 
  geom_pointrange() +  
  geom_hline(yintercept = 0, col="darkgrey", lty = 3, lwd=2) + 
  coord_flip() + theme_few() + ggtitle("Single main effect plot")

head(rdw_septic_group)
```

### Among sick animals (SNS and septic)

```{r}
variables = names(rdw_subset_sick)[(names(rdw_subset_sick) %in% c('days_of_hospitalization','discharge_status','patient_breed','sex','age','group','HCT','MCHC','RDW','anisocytosis','WBC','neut','lymph','mono','eos','baso','platelets','TPP','fib','RPR','creatinine'))]

rdw_subset_sick$surv = factor(rdw_subset_sick$survival_status)

getBetaAndSe <- function(covar){
  coefs <- glm(surv~scale(rdw_subset_sick[[covar]]),data=rdw_subset_sick,family='binomial') %>% 
    summary() %>% coef()
  data.frame(variable = covar, 
             estimate = coefs[2,1], 
             se = coefs[2,2], 
             p = coefs[2,4])
}

beta.table <- adply(t(covariates), 2, getBetaAndSe) %>% 
  arrange(estimate) %>% 
  mutate(low = estimate - 2*se, 
         high = estimate + 2*se,
         variable = reorder(variable, order(estimate)))

ggplot(beta.table, aes(y=estimate, x=variable, ymin=low, ymax=high, color = (p < 0.05))) + 
  geom_pointrange() +  
  geom_hline(yintercept = 0, col="darkgrey", lty = 3, lwd=2) + 
  coord_flip() + theme_few() + ggtitle("Single main effect plot")

head(rdw_subset_sick)
```

## Univariate analysis

### Among all \~ survival_status (survivor and non-survivor)

Run univariate correlation analysis for each possible covariate:

```{r}
rdw_log <- rdw_log %>%
  mutate (surv = ifelse(survival_status=='Survivor',1,0)) %>%
  mutate (surv = as.factor(surv))
# 1 means survival
# 0 means non-survival


fit.0 <- glm(surv~days_of_hospitalization, data = rdw_log, family = binomial(link=logit))
summary(fit.0)
#p-value 0.0249, estimate 0.11408

fit.1 <- glm(surv~discharge_status, data = rdw_log, family = binomial(link=logit))
summary(fit.1)
#this is garbage

fit.2 <- glm(surv~patient_breed, data = rdw_log, family = binomial(link=logit))
summary(fit.2)
#this is also garbage

fit.3 <- glm(surv~sex, data = rdw_log, family = binomial(link=logit))
summary(fit.3)
#females p-value 3.69e-15, estimate 1.93076

fit.4 <- glm(surv~age, data = rdw_log, family = binomial(link=logit))
summary(fit.4)
#garbage

fit.5 <- glm(surv~group, data = rdw_log, family = binomial(link=logit))
summary(fit.5)
#garbage

fit.6 <- glm(surv~HCT, data = rdw_log, family = binomial(link=logit))
summary(fit.6)
#garbage

fit.7 <- glm(surv~MCHC, data = rdw_log, family = binomial(link=logit))
summary(fit.7)
#p-value 0.0108, estimate 0.2653 

fit.8 <- glm(surv~RDW, data = rdw_log, family = binomial(link=logit))
summary(fit.8)
#nope

fit.9 <- glm(surv~anisocytosis, data = rdw_log, family = binomial(link=logit))
summary(fit.9)
#garbage

fit.10 <- glm(surv~WBC, data = rdw_log, family = binomial(link=logit))
summary(fit.10)
#crap

fit.11 <- glm(surv~neut, data = rdw_log, family = binomial(link=logit))
summary(fit.11)
#p-value 0.00458, estimate 1.757e-04

fit.12 <- glm(surv~lymph, data = rdw_log, family = binomial(link=logit))
summary(fit.12)
#poop

fit.13 <- glm(surv~mono, data = rdw_log, family = binomial(link=logit))
summary(fit.13)
#coco

fit.14 <- glm(surv~eos, data = rdw_log, family = binomial(link=logit))
summary(fit.14)
#nope

fit.15 <- glm(surv~baso, data = rdw_log, family = binomial(link=logit))
summary(fit.15)
#p value 0.0213, estimate 0.03647

fit.16 <- glm(surv~platelets, data = rdw_log, family = binomial(link=logit))
summary(fit.16)
#nope

fit.17 <- glm(surv~TPP, data = rdw_log, family = binomial(link=logit))
summary(fit.17)
#nope

fit.18 <- glm(surv~fib, data = rdw_log, family = binomial(link=logit))
summary(fit.18)
#p-value 0.000574, estimate -0.004173

fit.19 <- glm(surv~RPR, data = rdw_log, family = binomial(link=logit))
summary(fit.19)
#NO )))):

fit.20 <- glm(surv~creatinine, data = rdw_log, family = binomial(link=logit))
summary(fit.20)
#p-value 0.0324, estimate -0.09903
```

### Among sick \~ survival_status (survivor and non-survivor)

```{r}
rdw_subset_sick <- rdw_subset_sick %>%
  mutate (surv = ifelse(survival_status=='Survivor',1,0)) %>%
  mutate (surv = as.factor(surv))

fit.21 <- glm(surv~days_of_hospitalization, data = rdw_subset_sick, family = binomial(link=logit))
summary(fit.21)
#p-value 0.00152, estimate 0.19165

fit.22 <- glm(surv~discharge_status, data = rdw_subset_sick, family = binomial(link=logit))
summary(fit.22)
#trash

fit.23 <- glm(surv~patient_breed, data = rdw_subset_sick, family = binomial(link=logit))
summary(fit.23)
#trash

fit.24 <- glm(surv~sex, data = rdw_subset_sick, family = binomial(link=logit))
summary(fit.24)
#p-value female 1.88e-09, estimate 1.5215

fit.25 <- glm(surv~age, data = rdw_subset_sick, family = binomial(link=logit))
summary(fit.25)
#nope

fit.26 <- glm(surv~group, data = rdw_subset_sick, family = binomial(link=logit))
summary(fit.26)
#p-value SNS 5.79e-09, estimate 2.7556
#p-value septic 0.0639, estimate 0.4155

fit.27 <- glm(surv~HCT, data = rdw_subset_sick, family = binomial(link=logit))
summary(fit.27)
#nope

fit.28 <- glm(surv~MCHC, data = rdw_subset_sick, family = binomial(link=logit))
summary(fit.28)
#p-value 0.0119, estimate 0.2564

fit.29 <- glm(surv~RDW, data = rdw_subset_sick, family = binomial(link=logit))
summary(fit.29)
#nope )))):

fit.30 <- glm(surv~anisocytosis, data = rdw_subset_sick, family = binomial(link=logit))
summary(fit.30)
#nope

fit.31 <- glm(surv~WBC, data = rdw_subset_sick, family = binomial(link=logit))
summary(fit.31)
#nope

fit.32 <- glm(surv~neut, data = rdw_subset_sick, family = binomial(link=logit))
summary(fit.32)
#p-value 0.03326, estimate 1.226e-04

fit.33 <- glm(surv~lymph, data = rdw_subset_sick, family = binomial(link=logit))
summary(fit.33)
#nope

fit.34 <- glm(surv~mono, data = rdw_subset_sick, family = binomial(link=logit))
summary(fit.34)
#nope

fit.35 <- glm(surv~eos, data = rdw_subset_sick, family = binomial(link=logit))
summary(fit.35)
#nope

fit.36 <- glm(surv~baso, data = rdw_subset_sick, family = binomial(link=logit))
summary(fit.36)
#p-value 0.0498, estimate 0.03007

fit.37 <- glm(surv~platelets, data = rdw_subset_sick, family = binomial(link=logit))
summary(fit.37)
#nope

fit.38 <- glm(surv~TPP, data = rdw_subset_sick, family = binomial(link=logit))
summary(fit.38)
#caca

fit.39 <- glm(surv~fib, data = rdw_subset_sick, family = binomial(link=logit))
summary(fit.39)
#p-value 0.012, estimate -0.002960

fit.40 <- glm(surv~RPR, data = rdw_subset_sick, family = binomial(link=logit))
summary(fit.40)
#noooooooooo )))):

fit.41 <- glm(surv~creatinine, data = rdw_subset_sick, family = binomial(link=logit))
summary(fit.41)
#p-value 0.0734, estimate -0.08636 
```

### Among septic \~ survival_status (survivor and non-survivor)

```{r}
rdw_septic_group <- rdw_septic_group %>%
  mutate (surv = ifelse(survival_status=='Survivor',1,0)) %>%
  mutate (surv = as.factor(surv))

fit.42 <- glm(surv~days_of_hospitalization, data = rdw_septic_group, family = binomial(link=logit))
summary(fit.42)
#p-value 0.000362, estimate 0.24091

fit.43 <- glm(surv~discharge_status, data = rdw_septic_group, family = binomial(link=logit))
summary(fit.43)
#nope

fit.44 <- glm(surv~patient_breed, data = rdw_septic_group, family = binomial(link=logit))
summary(fit.44)
#nope

fit.45 <- glm(surv~sex, data = rdw_septic_group, family = binomial(link=logit))
summary(fit.45)
#nope

fit.46 <- glm(surv~age, data = rdw_septic_group, family = binomial(link=logit))
summary(fit.46)
#no

fit.47 <- glm(surv~HCT, data = rdw_septic_group, family = binomial(link=logit))
summary(fit.47)
#no

fit.48 <- glm(surv~MCHC, data = rdw_septic_group, family = binomial(link=logit))
summary(fit.48)
#estimate 0.2632, p value 0.0355

fit.49 <- glm(surv~RDW, data = rdw_septic_group, family = binomial(link=logit))
summary(fit.49)
#nope 

fit.50 <- glm(surv~anisocytosis, data = rdw_septic_group, family = binomial(link=logit))
summary(fit.50)
#nope

fit.51 <- glm(surv~WBC, data = rdw_septic_group, family = binomial(link=logit))
summary(fit.51)
#nope

fit.52 <- glm(surv~neut, data = rdw_septic_group, family = binomial(link=logit))
summary(fit.52)
#nope

fit.53 <- glm(surv~lymph, data = rdw_septic_group, family = binomial(link=logit))
summary(fit.53)
#nope

fit.54 <- glm(surv~mono, data = rdw_septic_group, family = binomial(link=logit))
summary(fit.54)
#nope

fit.55 <- glm(surv~eos, data = rdw_septic_group, family = binomial(link=logit))
summary(fit.55)
#nope

fit.56 <- glm(surv~baso, data = rdw_septic_group, family = binomial(link=logit))
summary(fit.56)
#nope

fit.57 <- glm(surv~platelets, data = rdw_septic_group, family = binomial(link=logit))
summary(fit.57)
#nope

fit.58 <- glm(surv~TPP, data = rdw_septic_group, family = binomial(link=logit))
summary(fit.58)
#nope

fit.59 <- glm(surv~fib, data = rdw_septic_group, family = binomial(link=logit))
summary(fit.59)
#nope

fit.60 <- glm(surv~RPR, data = rdw_septic_group, family = binomial(link=logit))
summary(fit.60)
#nope

fit.61 <- glm(surv~creatinine, data = rdw_septic_group, family = binomial(link=logit))
summary(fit.61)
#nope

```

### Among the ones that had blood culture done \~ survival

```{r}
rdw_blood_culture_done <- rdw_log %>%
  filter(blood_culture_done==1)%>%
  mutate (sepsis = ifelse(blood_culture_result=='1',1,0)) %>%
  mutate (sepsis = as.factor(sepsis))%>%
  mutate (surv = ifelse(survival_status=='Survivor',1,0)) %>%
  mutate (surv = as.factor(surv))

fit.62 <- glm(surv~days_of_hospitalization, data = rdw_blood_culture_done, family = binomial(link=logit))
summary(fit.62)
#p-value 0.00258, estimate 0.23610

fit.63 <- glm(surv~discharge_status, data = rdw_blood_culture_done, family = binomial(link=logit))
summary(fit.63)
#no

fit.64 <- glm(surv~patient_breed, data = rdw_blood_culture_done, family = binomial(link=logit))
summary(fit.64)
#no

fit.65 <- glm(surv~sex, data = rdw_blood_culture_done, family = binomial(link=logit))
summary(fit.65)
#p-value female 0.00442, estimate 1.0033

fit.66 <- glm(surv~age, data = rdw_blood_culture_done, family = binomial(link=logit))
summary(fit.66)
#no

fit.67 <- glm(surv~HCT, data = rdw_blood_culture_done, family = binomial(link=logit))
summary(fit.67)
#no

fit.68 <- glm(surv~MCHC, data = rdw_blood_culture_done, family = binomial(link=logit))
summary(fit.68)
#no

fit.69 <- glm(surv~RDW, data = rdw_blood_culture_done, family = binomial(link=logit))
summary(fit.69)
#nope ))):

fit.70 <- glm(surv~anisocytosis, data = rdw_blood_culture_done, family = binomial(link=logit))
summary(fit.70)
#no

fit.71 <- glm(surv~WBC, data = rdw_blood_culture_done, family = binomial(link=logit))
summary(fit.71)
#no

fit.72 <- glm(surv~neut, data = rdw_blood_culture_done, family = binomial(link=logit))
summary(fit.72)
#no

fit.73 <- glm(surv~lymph, data = rdw_blood_culture_done, family = binomial(link=logit))
summary(fit.73)
#no

fit.74 <- glm(surv~mono, data = rdw_blood_culture_done, family = binomial(link=logit))
summary(fit.74)
#no

fit.75 <- glm(surv~eos, data = rdw_blood_culture_done, family = binomial(link=logit))
summary(fit.75)
#no

fit.76 <- glm(surv~baso, data = rdw_blood_culture_done, family = binomial(link=logit))
summary(fit.76)
#no

fit.77 <- glm(surv~platelets, data = rdw_blood_culture_done, family = binomial(link=logit))
summary(fit.77)
#no

fit.78 <- glm(surv~TPP, data = rdw_blood_culture_done, family = binomial(link=logit))
summary(fit.78)
#no

fit.79 <- glm(surv~fib, data = rdw_blood_culture_done, family = binomial(link=logit))
summary(fit.79)
#p-value 0.039306, estimate -0.003185

fit.80 <- glm(surv~RPR, data = rdw_blood_culture_done, family = binomial(link=logit))
summary(fit.80)
#p-value 0.079790
# estimate -7.9501

fit.81 <- glm(surv~creatinine, data = rdw_blood_culture_done, family = binomial(link=logit))
summary(fit.81)
#no

```

### Among the ones that had blood culture done \~ sepsis (1=positive or 0=negative)

```{r}
rdw_blood_culture_done <- rdw_log %>%
  filter(blood_culture_done==1)%>%
  mutate (sepsis = ifelse(blood_culture_result=='1',1,0)) %>%
  mutate (sepsis = as.factor(sepsis))

fit.82 <- glm(sepsis~days_of_hospitalization, data = rdw_blood_culture_done, family = binomial(link=logit))
summary(fit.82)
#no

fit.83 <- glm(sepsis~discharge_status, data = rdw_blood_culture_done, family = binomial(link=logit))
summary(fit.83)
#ALIVE: p value 4.86e-06, estimate -1.3683
#EUTH: p value 0.00573, estimate 1.6196

fit.84 <- glm(sepsis~patient_breed, data = rdw_blood_culture_done, family = binomial(link=logit))
summary(fit.84)
#no

#rdw_blood_culture_done$sex <- relevel(rdw_blood_culture_done$sex, ref = "M")

fit.85 <- glm(sepsis~sex, data = rdw_blood_culture_done, family = binomial(link=logit))
summary(fit.85)
#females: p value 0.000774, estimate -1.2685

fit.86 <- glm(sepsis~age, data = rdw_blood_culture_done, family = binomial(link=logit))
summary(fit.86)
#no

fit.87 <- glm(sepsis~HCT, data = rdw_blood_culture_done, family = binomial(link=logit))
summary(fit.87)
#no

fit.88 <- glm(sepsis~MCHC, data = rdw_blood_culture_done, family = binomial(link=logit))
summary(fit.88)
#no

fit.89 <- glm(sepsis~RDW, data = rdw_blood_culture_done, family = binomial(link=logit))
summary(fit.89)
#nope ))):

fit.90 <- glm(sepsis~anisocytosis, data = rdw_blood_culture_done, family = binomial(link=logit))
summary(fit.90)
#no

fit.91 <- glm(sepsis~WBC, data = rdw_blood_culture_done, family = binomial(link=logit))
summary(fit.91)
#p value 0.031, estimate -0.0001620

fit.92 <- glm(sepsis~neut, data = rdw_blood_culture_done, family = binomial(link=logit))
summary(fit.92)
#p value 0.00547, estimate  -2.435e-04

fit.93 <- glm(sepsis~lymph, data = rdw_blood_culture_done, family = binomial(link=logit))
summary(fit.93)
#no

fit.94 <- glm(sepsis~mono, data = rdw_blood_culture_done, family = binomial(link=logit))
summary(fit.94)
#no

fit.95 <- glm(sepsis~eos, data = rdw_blood_culture_done, family = binomial(link=logit))
summary(fit.95)
#no

fit.96 <- glm(sepsis~baso, data = rdw_blood_culture_done, family = binomial(link=logit))
summary(fit.96)
#no

fit.97 <- glm(sepsis~platelets, data = rdw_blood_culture_done, family = binomial(link=logit))
summary(fit.97)
#no

fit.98 <- glm(sepsis~TPP, data = rdw_blood_culture_done, family = binomial(link=logit))
summary(fit.98)
#no

fit.99 <- glm(sepsis~fib, data = rdw_blood_culture_done, family = binomial(link=logit))
summary(fit.99)
#no

fit.100 <- glm(sepsis~RPR, data = rdw_blood_culture_done, family = binomial(link=logit))
summary(fit.100)
#no )))):

fit.101 <- glm(sepsis~creatinine, data = rdw_blood_culture_done, family = binomial(link=logit))
summary(fit.101)
#no

```

### Among all \~ sepsis (1=positive or 0=negative)

```{r}
rdw_all <- rdw_raw %>%
  mutate (sepsis = ifelse(group=='Septic',1,0)) %>%
  mutate (sepsis = as.factor(sepsis))


fit.102 <- glm(sepsis~days_of_hospitalization, data = rdw_all, family = binomial(link=logit))
summary(fit.102)
#estimate 0.04466, p value 0.0154

fit.103 <- glm(sepsis~discharge_status, data = rdw_all, family = binomial(link=logit))
summary(fit.103)
#alive estimate -1.5173, p value < 2e-16
#died estimate 2.6159, p value 0.00165
#euth estimate 3.4269, p value 8.21e-10

fit.104 <- glm(sepsis~patient_breed, data = rdw_all, family = binomial(link=logit))
summary(fit.104)
#no

fit.105 <- glm(sepsis~sex, data = rdw_all, family = binomial(link=logit))
summary(fit.105)
#female estimate -1.2272, p value 3.12e-10

fit.106 <- glm(sepsis~age, data = rdw_all, family = binomial(link=logit))
summary(fit.106)
#no

fit.107 <- glm(sepsis~HCT, data = rdw_all, family = binomial(link=logit))
summary(fit.107)
#no

fit.108 <- glm(sepsis~MCHC, data = rdw_all, family = binomial(link=logit))
summary(fit.108)
#no

fit.109 <- glm(sepsis~RDW, data = rdw_all, family = binomial(link=logit))
summary(fit.109)
#no
#estimate 0.004695, p value 0.974

fit.110 <- glm(sepsis~anisocytosis, data = rdw_all, family = binomial(link=logit))
summary(fit.110)
#no

fit.111 <- glm(sepsis~WBC, data = rdw_all, family = binomial(link=logit))
summary(fit.111)
# estimate -1.885e-04, p value 3.39e-05

fit.112 <- glm(sepsis~neut, data = rdw_all, family = binomial(link=logit))
summary(fit.112)
# estimate -2.521e-04, p value 5e-07

fit.113 <- glm(sepsis~lymph, data = rdw_all, family = binomial(link=logit))
summary(fit.113)
#estimate -0.0004911, p value 0.0454

fit.114 <- glm(sepsis~mono, data = rdw_all, family = binomial(link=logit))
summary(fit.114)
#estimate 0.0014373, p value 0.0111

fit.115 <- glm(sepsis~eos, data = rdw_all, family = binomial(link=logit))
summary(fit.115)
#no

fit.116 <- glm(sepsis~baso, data = rdw_all, family = binomial(link=logit))
summary(fit.116)
#estimate -0.05518, p value 2.33e-05

fit.117 <- glm(sepsis~platelets, data = rdw_all, family = binomial(link=logit))
summary(fit.117)
#estimate -0.006904, p value 0.00141

fit.118 <- glm(sepsis~TPP, data = rdw_all, family = binomial(link=logit))
summary(fit.118)
#no

fit.119 <- glm(sepsis~fib, data = rdw_all, family = binomial(link=logit))
summary(fit.119)
#estimate 0.004616, p value 2.91e-05

fit.120 <- glm(sepsis~RPR, data = rdw_all, family = binomial(link=logit))
summary(fit.120)
#estimate 12.6758, p value 0.000258

fit.121 <- glm(sepsis~creatinine, data = rdw_all, family = binomial(link=logit))
summary(fit.121)
#estimate 0.33784, p value 0.000177
```

### Among the sick \~ sepsis (1=positive or 0=negative)

```{r}

rdw_subset_sick <- rdw_subset_sick %>%
  mutate (sepsis = ifelse(group=='Septic',1,0)) %>%
  mutate (sepsis = as.factor(sepsis))

fit.122 <- glm(sepsis~days_of_hospitalization, data = rdw_subset_sick, family = binomial(link=logit))
summary(fit.122)
#no

fit.123 <- glm(sepsis~discharge_status, data = rdw_subset_sick, family = binomial(link=logit))
summary(fit.123)
#alive estimate -1.0508 , p value .60e-10
#died estimate 2.1494, p value 0.00986
#euth estimate 2.9604, p value 1.27e-07

fit.124 <- glm(sepsis~patient_breed, data = rdw_subset_sick, family = binomial(link=logit))
summary(fit.124)
#no

fit.125 <- glm(sepsis~sex, data = rdw_subset_sick, family = binomial(link=logit))
summary(fit.125)
#female estimate -0.7503, p value 0.000311

fit.126 <- glm(sepsis~age, data = rdw_subset_sick, family = binomial(link=logit))
summary(fit.126)
#no

fit.127 <- glm(sepsis~HCT, data = rdw_subset_sick, family = binomial(link=logit))
summary(fit.127)
#no

fit.128 <- glm(sepsis~MCHC, data = rdw_subset_sick, family = binomial(link=logit))
summary(fit.128)
#no

fit.129 <- glm(sepsis~RDW, data = rdw_subset_sick, family = binomial(link=logit))
summary(fit.129)
#no
#estimate -0.07826, p value 0.605

fit.120 <- glm(sepsis~anisocytosis, data = rdw_subset_sick, family = binomial(link=logit))
summary(fit.120)
#no

fit.131 <- glm(sepsis~WBC, data = rdw_subset_sick, family = binomial(link=logit))
summary(fit.131)
# estimate  -1.373e-04, p value 0.00143

fit.132 <- glm(sepsis~neut, data = rdw_subset_sick, family = binomial(link=logit))
summary(fit.132)
# estimate -1.936e-04, p value 5.68e-05

fit.133 <- glm(sepsis~lymph, data = rdw_subset_sick, family = binomial(link=logit))
summary(fit.133)
#no 

fit.134 <- glm(sepsis~mono, data = rdw_subset_sick, family = binomial(link=logit))
summary(fit.134)
#estimate 0.0013030, p value 0.0352

fit.135 <- glm(sepsis~eos, data = rdw_subset_sick, family = binomial(link=logit))
summary(fit.135)
#no

fit.136 <- glm(sepsis~baso, data = rdw_subset_sick, family = binomial(link=logit))
summary(fit.136)
#estimate -0.04973, p value 0.000117

fit.137 <- glm(sepsis~platelets, data = rdw_subset_sick, family = binomial(link=logit))
summary(fit.137)
#estimate -0.005284, p value 0.0145

fit.138 <- glm(sepsis~TPP, data = rdw_subset_sick, family = binomial(link=logit))
summary(fit.138)
#no

fit.139 <- glm(sepsis~fib, data = rdw_subset_sick, family = binomial(link=logit))
summary(fit.139)
#estimate 0.002892, p value 0.00719

fit.140 <- glm(sepsis~RPR, data = rdw_subset_sick, family = binomial(link=logit))
summary(fit.140)
#estimate 9.5720, p value 0.00535

fit.141 <- glm(sepsis~creatinine, data = rdw_subset_sick, family = binomial(link=logit))
summary(fit.141)
#estimate 0.27167, p value 0.00118
```

## Trying some models

### Survival within all

```{r}
model1.1 <- glm(surv~MCHC+RDW+neut+baso+fib+RPR+creatinine, data = rdw_log, family = binomial(link=logit))

model1.2 <- glm(surv~RPR+MCHC+neut+baso+fib+creatinine, data = rdw_log, family = binomial(link=logit))
stepAIC(model1.1)

lroc(model1.1,grid=FALSE,line.col = "red",auc.coords=c(.25,.1))
lroc(model1.2,grid=FALSE,line.col = "red",auc.coords=c(.25,.1))
lroc(model1.3,grid=FALSE,line.col = "red",auc.coords=c(.25,.1))

#the best model here was model1.1
```

```{r}
glm_out_1.1 <- glm(formula = sepsis ~ RPR, family = binomial(link = logit), data = rdw_all)%>%
  augment() %>%
  mutate(model='RPR')

glm_out_rpr %>%
  roc_curve(event_level = 'second', truth = sepsis, .fitted)%>%
  ggplot(aes(x = 1 - specificity, y = sensitivity))+
  geom_line(size = 1.1,color='red') +
  geom_abline(slope = 1, intercept = 0, size = 0.4) +
  coord_fixed() +
  theme_cowplot()

glm_out_rpr %>%
  roc_auc(event_level = 'second', truth = sepsis, .fitted)

```

### Survival within sick

```{r}
model2.1 <- glm(formula = surv ~ MCHC + RDW + neut + baso + fib + RPR, family = binomial(link = logit), data = rdw_subset_sick)

summary(model2.1)

lroc(model2.1,grid=FALSE,line.col = "red",auc.coords=c(.25,.1))

stepAIC(model2.1,grid=FALSE,line.col = "red",auc.coords=c(.25,.1))

model2.2 <- glm(formula = surv ~ MCHC + RPR + neut + baso + fib, family = binomial(link = logit), data = rdw_subset_sick)

lroc(model2.2,grid=FALSE,line.col = "red",auc.coords=c(.25,.1))
```

### Survival within septic

```{r}
model3.1 <- glm(formula = surv ~ MCHC+RPR+RDW, family = binomial(link = logit), data = rdw_septic_group)

summary(model3.1)

stepAIC(model3.1)

lroc(model3.1,grid=FALSE,line.col = "red",auc.coords=c(.25,.1))

model3.2 <- glm(formula = surv ~ MCHC, family = binomial(link = logit), data = rdw_septic_group)

lroc(model3.2,grid=FALSE,line.col = "red",auc.coords=c(.25,.1))

```

### Survival within the ones that had blood culture done

```{r}
model3 <- glm(formula = sepsis ~ days_of_hospitalization+sex+RDW+fib+RPR, family = binomial(link = logit), data = rdw_blood_culture_done)

summary(model3)

```

### Sepsis within the ones that had blood culture done

```{r}
model4.1<- glm(formula = sepsis ~ RPR +  RDW + WBC + neut, family = binomial(link = logit), data = rdw_blood_culture_done)

summary(model4.1)

lroc(model4.1,grid=FALSE,line.col = "red",auc.coords=c(.25,.1))


model4.2 <- glm(formula = sepsis ~ WBC + neut, family = binomial(link = logit), data = rdw_blood_culture_done)

summary(model4.2)

lroc(model4.2,grid=FALSE,line.col = "red",auc.coords=c(.25,.1))
```

### Sepsis within septic group

```{r}
model5.1 <- glm(formula = sepsis ~ RPR +  RDW + WBC + neut, family = binomial(link = logit), data = rdw_blood_culture_done)

summary(model5.1)

lroc(model5.1,grid=FALSE,line.col = "red",auc.coords=c(.25,.1))


model5.2 <- glm(formula = sepsis ~ WBC + neut, family = binomial(link = logit), data = rdw_blood_culture_done)

summary(model5.2)

lroc(model5.2,grid=FALSE,line.col = "red",auc.coords=c(.25,.1))
```

### Sepsis within all

```{r}
model6<- glm(formula = sepsis ~ RPR + WBC + neut + lymph + mono + baso + platelets + creatinine, family = binomial(link = logit), data = rdw_all)

summary(model6)

lroc(model6,grid=FALSE,line.col = "red",auc.coords=c(.25,.1))

stepAIC(model6)

best.model <- glm(formula = sepsis ~ RPR + WBC + neut + mono + baso + creatinine, family = binomial(link = logit), data = rdw_all)

summary(best.model)

lroc(best.model,grid=FALSE,line.col = "red",auc.coords=c(.25,.1))

```

## Univariate ROCs

### RPR

```{r}
rdw_all <- rdw_all%>%
  mutate(sepsis2=as.factor(sepsis))

model.rpr <- glm(formula = sepsis ~ RPR, family = binomial(link = logit), data = rdw_all)

lroc.rpr <- lroc(model.rpr,grid=FALSE,line.col = "red",auc.coords=c(.25,.1))

#rocobj <- roc(rdw_all$sepsis2,rdw_all$RPR)

#roc.info <- roc(rdw_all$sepsis2, model.rpr$fitted.values,legacy.axes=TRUE)


#plot(x=rdw_all$RPR,y=rdw_all$sepsis)
#model.rpr <- glm(formula = sepsis ~ RPR, family = binomial(link = logit), data = rdw_all)
#lines(rdw_all$sepsis,model.rpr$fitted.values)

```

```{r}
glm_out_rpr <- glm(formula = sepsis ~ RPR, family = binomial(link = logit), data = rdw_all)%>%
  augment() %>%
  mutate(model='RPR')

glm_out_rpr %>%
  roc_curve(event_level = 'second', truth = sepsis, .fitted)%>%
  ggplot(aes(x = 1 - specificity, y = sensitivity))+
  geom_line(size = 1.1,color='red') +
  geom_abline(slope = 1, intercept = 0, size = 0.4) +
  coord_fixed() +
  theme_cowplot() # + 
  #geom_smooth()

glm_out_rpr %>%
  roc_auc(event_level = 'second', truth = sepsis, .fitted)

library(cutpointr)

cp <- cutpointr(rdw_all,RPR, sepsis, method = maximize_metric, metric = youden)
summary(cp)
```

### WBC

```{r}
model.wbc <- glm(formula = sepsis ~ WBC, family = binomial(link = logit), data = rdw_all)

lroc(model.wbc,grid=FALSE,line.col = "red",auc.coords=c(.25,.1))
```

### Neutrophils

```{r}
model.neut <- glm(formula = sepsis ~ neut, family = binomial(link = logit), data = rdw_all)

lroc(model.neut,grid=FALSE,line.col = "red",auc.coords=c(.25,.1))
```

### Lymphocytes

```{r}
model.lymph <- glm(formula = sepsis ~ lymph, family = binomial(link = logit), data = rdw_all)

lroc(model.lymph,grid=FALSE,line.col = "red",auc.coords=c(.25,.1))
```

### Monocytes

```{r}
model.mono <- glm(formula = sepsis ~ mono, family = binomial(link = logit), data = rdw_all)

lroc(model.mono,grid=FALSE,line.col = "red",auc.coords=c(.25,.1))
```

### Basophils

```{r}
model.baso <- glm(formula = sepsis ~ baso, family = binomial(link = logit), data = rdw_all)

lroc(model.baso,grid=FALSE,line.col = "red",auc.coords=c(.25,.1))
```

### Creatinine

```{r}
model.creatinine <- glm(formula = sepsis ~ creatinine, family = binomial(link = logit), data = rdw_all)

lroc(model.creatinine,grid=FALSE,line.col = "red",auc.coords=c(.25,.1))

lroc7<-lroc(model.creatinine, cex=0.8,grid=FALSE,percent=TRUE, line.col="grey", lty=1,lwd=1.3)$auc

```

```{r}
lroc(best.model,cex=0.8,line.col="red",grid=FALSE, lwd=0.8)$auc
lroc1<-lroc(model.rpr, add=TRUE,cex=0.8,grid=FALSE,line.col="black", lty=1,lwd=1.3)$auc
lroc2<-lroc(model.wbc, add=TRUE,cex=0.8,grid=FALSE,line.col="blue", lty=1,lwd=1.3)$auc
lroc3<-lroc(model.neut, add=TRUE,cex=0.8,grid=FALSE,line.col="darkgreen", lty=1,lwd=1.3)$auc
lroc4<-lroc(model.lymph, add=TRUE,cex=0.8,grid=FALSE,line.col="orange", lty=1,lwd=1.3)$auc
lroc5<-lroc(model.mono, add=TRUE,cex=0.8,grid=FALSE, line.col="purple", lty=1,lwd=1.3)$auc
lroc6<-lroc(model.baso, add=TRUE,cex=0.8,grid=FALSE, line.col="grey", lty=1,lwd=1.3)$auc
lroc7<-lroc(model.creatinine, add=TRUE,cex=0.8,grid=FALSE,  line.col="pink", lty=1,lwd=1.3)$auc

```

# Overall ROC curve

```{r}
glm_out1 <- glm(formula = sepsis ~ RPR, family = binomial(link = logit), data = rdw_all)%>%
  augment() %>%
  mutate(model='RPR (AUC = 63.2%)')

glm_out2 <- glm(formula = sepsis ~ WBC, family = binomial(link = logit), data = rdw_all)%>%
  augment()%>%
  mutate(model='WBC (AUC = 70.1%)')

glm_out3 <- glm(formula = sepsis ~ neut, family = binomial(link = logit), data = rdw_all)%>%
  augment() %>%
  mutate(model='Neutrophils (AUC = 73.4%)')

glm_out4 <- glm(formula = sepsis ~ lymph, family = binomial(link = logit), data = rdw_all)%>%
  augment() %>%
  mutate(model='Lymphocytes (AUC = 62.8%)')

glm_out5 <- glm(formula = sepsis ~ mono, family = binomial(link = logit), data = rdw_all)%>%
  augment() %>%
  mutate(model='Monocytes (AUC = 55.5%)') 

glm_out6 <- glm(formula = sepsis ~ baso, family = binomial(link = logit), data = rdw_all)%>%
  augment() %>%
  mutate(model='Basophils (AUC = 75.2%)')

glm_out7 <- glm(formula = sepsis ~ creatinine, family = binomial(link = logit), data = rdw_all)%>%
  augment() %>%
  mutate(model='Creatinine (AUC = 66%)')

glm_out8 <- glm(formula = sepsis ~ RPR + WBC + neut + lymph + mono + baso + platelets + creatinine, family = binomial(link = logit), data = rdw_all)%>%
  augment() %>%
  mutate(model='Final model (AUC = 82.1%)')

glm_out <- bind_rows (glm_out1,glm_out2,glm_out3,glm_out4,glm_out5,glm_out6, glm_out7,glm_out8)

glm_out %>%
  group_by(model)%>%
  roc_curve(event_level = 'second', truth = sepsis,.fitted)%>%
  ggplot(aes(x = 1 - specificity, y = sensitivity, color = model))+
  geom_line(size = 1.3) +
  geom_abline(slope = 1.5, intercept = 0, size = 0.4) +
  coord_fixed() +
  theme_cowplot()+
  scale_fill_discrete(breaks=c('final_model','baso','neut','WBC','creatinine','RPR','lymph','mono' ))
```

# Survival analysis

```{r}
rdw_log <- rdw_log %>%
  mutate (death = ifelse(survival_status == 'Survivor',0,1)) %>%
  mutate (death = as.numeric(death))

rdw_surv <- rdw_log %>%
  filter(!patient_number %in% c('687982','698921','838471'))

surv_1 = Surv(rdw_surv$days_of_hospitalization, rdw_surv$death)

KMfit = survfit(surv_1 ~ group, data = rdw_surv, stype=1, ctype=1)
KMfit

summary(KMfit)

```

```{r}
plot(KMfit, mark.time=TRUE, xlab='Survival time (in days)', ylab='Survival by group',ylim=c(0.4,1), conf.int=FALSE, col=c('green','red','blue'), cex=0.3)
legend("bottomright",1,c('Healthy','SNS','Septic'), col=c('green','blue','red'),lty=1) 


surv_diff <- survdiff(Surv(days_of_hospitalization, death) ~ group, data = rdw_surv)

surv_diff
```

```{r}
rdw_surv <- rdw_surv %>%
  mutate (death = ifelse(survival_status == 'Survivor',0,1)) %>%
  mutate (death = as.numeric(death))%>%
  mutate (RPR_binary = ifelse(RPR >= '0.0928',1,0))

surv_2 = Surv(rdw_surv$days_of_hospitalization, rdw_surv$death)

KMfit2 = survfit(surv_2 ~ RPR_binary, data = rdw_surv, stype=1, ctype=1)
KMfit2

summary(KMfit2)

summary(KMfit2)$table

```

```{r}
plot(KMfit2, mark.time=TRUE, xlab='Days of hospitalization', ylab='Survival rate by RPR level',ylim=c(0.4,1), xlim=c(0,35),conf.int=FALSE, col=c('red','blue'), cex=0.3,lty=c(1,1))
legend("bottomright",1,c('RPR < 0.0928','RPR  0.0928'), col=c('red','blue'),lty=1) 

surv_diff2 <- survdiff(Surv(days_of_hospitalization, death) ~ RPR_binary, data = rdw_surv)

surv_diff2


```

# Medians based on RPR cutoff

## Creating subsets

```{r}
RPR_above_cutoff <- rdw_all %>%
  filter(RPR_binary %in% c('1'))
dim(RPR_above_cutoff)

RPR_below_cutoff <- rdw_all %>%
  filter(RPR_binary %in% c('0'))
dim(RPR_below_cutoff)
dim(rdw_all)
```

## Wilcoxon test

```{r}
wilcox.test(baso~RPR_binary,data=rdw_all) 
wilcox.test(creatinine~RPR_binary,data=rdw_all) 
wilcox.test(CRT~RPR_binary,data=rdw_all) 
wilcox.test(eos~RPR_binary,data=rdw_all) 
wilcox.test(glu~RPR_binary,data=rdw_all) 
wilcox.test(P~RPR_binary,data=rdw_all) 
wilcox.test(HCT~RPR_binary,data=rdw_all) 
wilcox.test(Hg~RPR_binary,data=rdw_all) 
wilcox.test(lactate~RPR_binary,data=rdw_all) 
wilcox.test(lymph~RPR_binary,data=rdw_all) 
wilcox.test(MCH~RPR_binary,data=rdw_all) 
wilcox.test(MCHC~RPR_binary,data=rdw_all) 
wilcox.test(MCV~RPR_binary,data=rdw_all) 
wilcox.test(mono~RPR_binary,data=rdw_all) 
wilcox.test(neut~RPR_binary,data=rdw_all) 
wilcox.test(platelets~RPR_binary,data=rdw_all) 
wilcox.test(RBC~RPR_binary,data=rdw_all) 
wilcox.test(RDW~RPR_binary,data=rdw_all) 
wilcox.test(R~RPR_binary,data=rdw_all) 
wilcox.test(RPR~RPR_binary,data=rdw_all) 
wilcox.test(sepsis_score~RPR_binary,data=rdw_all) 
wilcox.test(T~RPR_binary,data=rdw_all) 
wilcox.test(WBC~RPR_binary,data=rdw_all) 
```

## Medians above cutoff

```{r}
RPR_above_cutoff$P <- as.numeric(RPR_above_cutoff$P)
RPR_above_cutoff$R <- as.numeric(RPR_above_cutoff$R)
RPR_above_cutoff$creatinine <- as.numeric(RPR_above_cutoff$creatinine)
RPR_above_cutoff$CRT <- as.numeric(RPR_above_cutoff$CRT)


MedianCI(RPR_above_cutoff$baso, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_above_cutoff$creatinine, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_above_cutoff$CRT, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_above_cutoff$eos, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_above_cutoff$glu, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_above_cutoff$P, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_above_cutoff$HCT, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_above_cutoff$Hg, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_above_cutoff$lactate, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_above_cutoff$lymph, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_above_cutoff$MCH, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_above_cutoff$MCHC, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_above_cutoff$MCV, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_above_cutoff$mono, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_above_cutoff$neut, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_above_cutoff$platelets, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_above_cutoff$RBC, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_above_cutoff$RDW, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_above_cutoff$R, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_above_cutoff$RPR, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_above_cutoff$sepsis_score, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_above_cutoff$T, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_above_cutoff$WBC, conf.level=0.95, na.rm=TRUE)



```

## Medians below cutoff

```{r}
RPR_below_cutoff$P <- as.numeric(RPR_below_cutoff$P)
RPR_below_cutoff$R <- as.numeric(RPR_below_cutoff$R)



MedianCI(RPR_below_cutoff$baso, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_below_cutoff$creatinine, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_below_cutoff$CRT, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_below_cutoff$eos, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_below_cutoff$glu, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_below_cutoff$P, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_below_cutoff$HCT, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_below_cutoff$Hg, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_below_cutoff$lactate, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_below_cutoff$lymph, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_below_cutoff$MCH, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_below_cutoff$MCHC, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_below_cutoff$MCV, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_below_cutoff$mono, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_below_cutoff$neut, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_below_cutoff$platelets, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_below_cutoff$RBC, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_below_cutoff$RDW, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_below_cutoff$R, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_below_cutoff$RPR, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_below_cutoff$sepsis_score, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_below_cutoff$T, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_below_cutoff$WBC, conf.level=0.95, na.rm=TRUE)

```

# Medians based on study groups

```{r}
MedianCI(rdw_all$baso, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_all$creatinine, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_all$CRT, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_all$eos, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_all$glu, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_all$P, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_all$HCT, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_all$Hg, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_all$lactate, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_all$lymph, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_all$MCH, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_all$MCHC, conf.level=0.95, na.rm=TRUE)
MedianCI(RPR_below_cutoff$MCV, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_all$mono, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_all$neut, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_all$platelets, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_all$RBC, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_all$RDW, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_all$R, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_all$RPR, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_all$sepsis_score, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_all$T, conf.level=0.95, na.rm=TRUE)
MedianCI(rdw_all$WBC, conf.level=0.95, na.rm=TRUE)

```
